---
title: "p8105_hw3_jz3571"
author: "Junyan Zhu"
date: "2022-10-16"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(patchwork)
library(ggplot2)

```

# Problem 1

```{r}
library(p8105.datasets)
data("instacart")
summary(instacart)
insta_df = instacart
```

## How many aisles are there and which aisles are the most items ordered from?
```{r}
num_aisle = insta_df %>% 
  group_by(aisle) %>% 
  summarise(total_num = n()) %>% 
  arrange(desc(total_num))

nrow(num_aisle)

names(which.max(table(insta_df$aisle)))

```

```{r}
num_aisle %>% 
  filter(total_num > 10000) %>% 
  ggplot() +
  geom_bar(aes(x = total_num, y = aisle), stat = "identity") +
  labs(title = "Number of Orders for Different Aisles",
       x = "Type of Aisel",
       y = "Number of Order") +
  theme_classic()
```

```{r}
three_pop_df = insta_df %>% 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarise(count = n()) %>% 
  mutate(rank = min_rank(desc(count))) %>% 
  filter(rank <= 3) %>% 
  arrange(aisle, rank)

three_pop_df
```

```{r}
mean_hr_df = insta_df %>% 
  filter(product_name %in% c("Pink Lady Apples", 'Coffee Ice Cream')) %>% 
  select(order_dow, order_hour_of_day, product_name) %>% 
  group_by(order_dow, product_name) %>% 
  summarise(mean(order_hour_of_day)) %>% 
  mutate(order_dow = recode(order_dow,
                            '0' = 'Sunday',
                            '1' = 'Monday',
                            '2' = 'Tuesday',
                            '3' = 'Wednesday',
                            '4' = 'Thursday',
                            '5' = 'Friday',
                            '6' = 'Saturday'))

mean_hr_df
```

# Problem 2

```{r}
acc_df = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names()
```

## Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
tidy_acc_df = acc_df %>% 
  mutate(weekdays = ifelse(day == "Saturday" | day =="Sunday", 'weekend', 'weekday')) %>% 
  mutate(activity_daysum = rowSums(.[4:1443]))
```

There are total 35 days of observations and 1443 variables exist, including week, day, day-id, and 1440 variables for each minute in a day. The new created weekdays variable specifies whether a recorded day belongs to weekday or weekend, and activity_daysum variable reflects the total activity count for a specific day by summing all 1440 counts for a day.

## Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
tidy_acc_df %>%  
  select(week, day, activity_daysum) %>% 
  arrange(week, day) %>% 
  knitr::kable()
```
There is no obvious pattern if we extract the table for total activity acount for each day. Let's try grouping the activity acount by specific day in the week.

```{r}
tidy_acc_df %>% 
  group_by(day) %>% 
  summarise(mean_acitivity = mean(activity_daysum)) %>% 
  knitr::kable()
```
There is still no obvious patter showed on the mean of activity for each specific day in the week. Wednesday, Thursday and Friday has higher activity acount than other days.
Let's try grouping by weekdays and weekends.

```{r}
tidy_acc_df %>% 
  group_by(weekdays) %>% 
  summarise(mean_activity = mean(activity_daysum)) %>% 
  knitr::kable()
```
Now we can see that weekdays have higher mean value of daily activity count than weekends.

## Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
plot_df = tidy_acc_df %>% 
  pivot_longer(activity_1:activity_1440,
               names_to = "minute",
               values_to = 'activity',
               names_prefix = "activity_") %>% 
  mutate(minute = as.integer(minute)) %>% 
  ggplot(aes(x = minute , y = activity, color = day)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  labs(title = "24-Hour activity count for each day",
       x = "Minute",
       y= 'Activity count') +
  scale_color_hue(name = "Day") +
  theme(plot.title = element_text(hjust = 0.5))
  
plot_df
```
We can see from the table the activity count is lowest during sleeping hours (0th - 400th minute), and has peaks at noon (around 700th minute) and in the evening (around 1250th min).

